Binary Image Transition State Search \\ Supplementary Information

### 2022-03-04 manuscript.tex
SI interpolation section added. Removed zeroeigen failure discussion

\topic However, a large range of landscapes prove challenging or impossible to explore via these methods.
One key problem arises from the push towards larger and more complex systems \cite{Trefethen2013,Shalf2020,Alexander2020}, resulting in the need to develop algorithms that are more memory and computationally efficient, and can incorporate optimisation strategies such as on-the-fly adaptive remeshing and coarse-graining.
Current double-ended methods require a chain of tens of states to gain an accurate approximation, and have issues with remeshing resulting in a different discretisation for each state.
Single-ended methods are not well suited for identifying specific pathways and can spend a large amount of time searching for undesired transition states.
Another major challenge in studying complex energy landscapes relates to the presence of locally flat or discontinous regions.
Both of which can be relevent when considering contact forces for macroscopic objects \temp{[cite]}, or hard-body interactions in atomistic systems \temp{[cite]}.
Flat zero-modes in the landscape pose issues for single-ended search methods that rely only upon local information.
Specialist treatment can sometimes be used such as in the case of global rotation and translation \temp{\cite{Wales2004}}, but they are thwarted by local zero-modes.
Meanwhile, current methods cannot typically be applied in the case of discontinuous potentials, or where the gradient is prohibitively expensive to compute, because they require continuous, differentiable optimisation functions.

@@

%\section{Zero-curvature modes}
\topic The final challenges we will address are those related to complex landscapes.
The first is the presence of flat regions in the landscape.
\Cref{fig:flatdiscontinuous}a shows BITSS applied to a 2D landscape with two such regions (i \& ii) that are flat in the $x$-direction.
We see BITSS is able to successfully converge past these flat regions, even with one very close to the transition state (ii).
In these regions there are no driving forces due to the potential and the energy constraint, which use purely local information about the gradient.
However, the distance constraint continues to pull the states together, preventing them from getting stuck.
When only a single state has a zero-gradient mode then the other is likely to slide down the potential slightly (iii), but the two states still remain either side of the dividing ridge and so the result is unaffected.
\temp{Indeed, the only circumstance under which a zero-mode could cause BITSS to fail is if it is positioned at the transition state in the direction of the pathway.
In this case there is no single transition state, instead the minimum energy pathway would have a flat section at the top of its energy profile.}

>>>>

\topic However, a large range of landscapes prove challenging or impossible to explore via these methods.
One key problem arises from the push towards larger and more complex systems \cite{Trefethen2013,Shalf2020,Alexander2020}, resulting in the need to develop algorithms that are more memory and computationally efficient, and can incorporate optimisation strategies such as on-the-fly adaptive remeshing and coarse-graining.
Current double-ended methods require a chain of tens of states to gain an accurate approximation, and have issues with remeshing resulting in a different discretisation for each state.
Single-ended methods are not well suited for identifying specific pathways and can spend a large amount of time searching for undesired transition states.
Another major challenge in studying complex energy landscapes relates to the presence of locally flat or discontinous regions.
Both of which can be relevent when considering contact forces for macroscopic objects \temp{[cite]}, or hard-body interactions in atomistic systems \temp{[cite]}.
Flat zero-modes in the landscape pose issues for single-ended search methods that rely only upon local information.
Specialist treatment can sometimes be used such as in the case of global rotation and translation \cite{Page1988}, but they are thwarted by local zero-modes.
Meanwhile, current methods cannot typically be applied in the case of discontinuous potentials, or where the gradient is prohibitively expensive to compute, because they require continuous, differentiable optimisation functions.

@@

%\section{Zero-curvature modes}
\topic The final challenges we will address are those related to complex landscapes.
The first is the presence of flat regions in the landscape.
\Cref{fig:flatdiscontinuous}a shows BITSS applied to a 2D landscape with two such regions (i \& ii) that are flat in the $x$-direction.
We see BITSS is able to successfully converge past these flat regions, even with one very close to the transition state (ii).
In these regions there are no driving forces due to the potential and the energy constraint, which use purely local information about the gradient.
However, the distance constraint continues to pull the states together, preventing them from getting stuck.
When only a single state has a zero-gradient mode then the other is likely to slide down the potential slightly (iii), but the two states still remain either side of the dividing ridge and so the result is unaffected.

### 2022-03-04 si/supplementary_information.tex
SI interpolation section added. Removed zeroeigen failure discussion

\newpage
\section{Additional results: DNEB speed comparison}
\Cref{fig:DNEBspeedtest} shows the results for the speed test in comparison with DNEB.
Much the same as the string method, the slope of BITSS is greater than DNEB with 10 or 20 states, indicating a faster convergence.
Furthermore, with the exception of the seven-particle cluster, DNEB is also much less accurate at approximating the transition state, even with a large number of images.
So it will require a secondary method to accurately locate the transition state, while BITSS does not.

@@

\section{Additional results: Seven-particle cluster comparison}
\Cref{fig:clusterdgraphs} shows the comparison between the energies of the minima and transition states for the seven-particle cluster.
Notably the energies of the minima are largely unchanged, but the energies of the transition states are found to be slightly higher when using the discontinuous potential.
This indicates that the particles in the Lennard-Jones cluster cut the corner slightly as they transition, whereas this is not possible using the discontinuous potential, resulting in higher energies.

>>>>

\section{Details for the adaptive discretisation tests}
Here we provide the details for the interpolations and mapping involved in the two examples demonstrating the feasibility of using an adaptive discretisation method.
For the cylindrical buckling example with a changing mesh, the resolution is refined each time the separation between the two states is halved. 
This involves the number of triangles around the circumference of the cylinder increasing along the sequence, $40 \rightarrow 60 \rightarrow 80 \rightarrow 100$.
The positions of the nodes on the new grid, $\{\bm{n}_i\}$, are determined by linear interpolating from the previous grid $\{\bm{p}_i\}$, using the positions of the unbuckled meshes, $\{\bm{n'}_i\}$ and $\{\bm{p'}_i\}$.
For each node of the new grid, $\bm{n'}_i$, the triangle that contains it is first identified, which we will denote $\{\bm{p'}_1,\bm{p'}_2,\bm{p'}_3\}$, and the barycentric coordinates of the point are computed, $\{\lambda_1,\lambda_2,\lambda_3\}$.
The new position is then given by $\bm{n}_i = \lambda_1 \bm{p}_1 + \lambda_2 \bm{p}_2 + \lambda_3 \bm{p}_3$.

In the example with different resolutions for the two states, the distance is obtained by mapping the phase field from the higher resolution grid to the low resolution grid, and then using the 2-norm.
Because a square grid is used with a resolution ratio of two, the mapping involves averaging each 2x2 block to a single point.
Finally, the gradient with respect to each point must be mapped back to the higher-resolution grid, which is done by assigning a quarter of each component back to its the original four points.

\section{Additional results: DNEB speed comparison}
\Cref{fig:DNEBspeedtest} shows the results for the speed test in comparison with DNEB.
Much the same as the string method, the slope of BITSS is greater than DNEB with 10 or 20 states, indicating a faster convergence.
Furthermore, with the exception of the seven-particle cluster, DNEB is also much less accurate at approximating the transition state, even with a large number of images.
So it will require a secondary method to accurately locate the transition state, while BITSS does not.

@@

\newpage
\section{Additional results: Seven-particle cluster comparison}
\Cref{fig:clusterdgraphs} shows the comparison between the energies of the minima and transition states for the seven-particle cluster.
Notably the energies of the minima are largely unchanged, but the energies of the transition states are found to be slightly higher when using the discontinuous potential.
This indicates that the particles in the Lennard-Jones cluster cut the corner slightly as they transition, whereas this is not possible using the discontinuous potential, resulting in higher energies.
